### 2022-06-27 main/chapter_2ljets.tex
begin modelling section; add hist* and pyhf references

\FloatBarrier
\section{Modelling}
% TODO backgorund samples
% TODO matrix method fakes
% TODO GMSB reweighting
% TODO off-shell plus/minus split
% TODO generator filtering

@@

\begin{table}[tp]
\centering
\begin{tabular}{lcccccc}
& Prompt (stat)
& mMC
& mMM
& mMean
& \quad\underline{mMax}\hphantom{foo}
\\[0.5em]
CR-Z & $143.6 \pm 4.6$
& $0.1^{+0.2}_{-0.1}$
& $0.01^{+3.2}_{-0.0}$
& $0.02^{+0.14}_{-0.02}$
& $0.1^{+3.2}_{-0.1}$
\\[0.2em]
VR-Int & $18.7 \pm 0.9$
& $0.2^{+0.4}_{-0.2}$
& $0.01^{+0.46}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.2^{+0.5}_{-0.2}$
\\[0.2em]
VR-High & $19.7 \pm 0.7$
& $0.04^{+0.08}_{-0.04}$
& $0.01^{+0.73}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.04^{+0.73}_{-0.04}$
\\[0.2em]
VR-High-1J & $51.9 \pm 1.9$
& $0.003^{+0.005}_{-0.003}$
& $0.01^{+0.59}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.59}_{-0.02}$
\\[0.2em]
SR-Low-b & $15.4 \pm 1.0$
& $0.001^{+0.001}_{-0.001}$
& $0.01^{+0.11}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.17}_{-0.02}$
\\[0.2em]
SR-Int-a & $22.3 \pm 0.8$
& $0^{+0}_{-0}$
& $0.01^{+0.72}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.72}_{-0.02}$
\\[0.2em]
SR-Int-b & $11.4 \pm 0.5$
& $0.002^{+0.003}_{-0.002}$
& $0.01^{+0.43}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.43}_{-0.02}$
\\[0.2em]
SR-High-8-a & $2.4 \pm 0.2$
& $0^{+0}_{-0}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\[0.2em]
SR-High-8-b & $2.4 \pm 0.3$
& $0^{+0}_{-0}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\[0.2em]
SR-High-1J & $1.0 \pm 0.1$
& $0.004^{+0.007}_{-0.004}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\[0.2em]
SR-$\llbb$ & $0.7 \pm 0.1$
& $0^{+0}_{-0}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\
\end{tabular}
\caption[%
Replacement fake estimates with up and down variations
]{%
Replacement fake estimates with up and down variations.
Those variations cover statistical and systematic sources.
The method `mMax' is used for the $\twoljets$-electroweak analysis;
as described in the text, it uses the largest values of central, statistical
and positive systematic variations from other methods, and a down variation
of zero.
}
\label{tab:2ljets_fnp_estimates}
\end{table}

\section{Variations}
% "systematic"
Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
Figure~\ref{fig:2ljets_low_crz_pre_post}
Figure~\ref{fig:2ljets_vrlow_alt}
Figure~\ref{fig:2ljets_low_vrlow_pre_post}

@@

% extrapolation

The total theory variation is a combination of seven components representing
\TODO{what are they?}.
A relative uncertainty due to each, $r$, is evaluated at each benchmark
point $\vec y_i$, and extrapolated elsewhere as an exponentially weighted
combination
\begin{equation}
\label{eqn:2ljets_signal_extrap}
r(\vec x) = \frac{
\sum_i r_i e^{-d(\vec x, \vec y_i)}
}{
\sum_i e^{-d(\vec x, \vec y_i)}
}
\end{equation}
in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
physical understanding into the extrapolation.

>>>>

\FloatBarrier
\section{Modelling}

So far, we have seen many histograms of background samples overlaid with
observed data.
Those backgrounds may or may not have appeared to predicted the data well, but
no conclusions can be drawn until uncertainties are assigned to the background
and signal modelling.

To interpret the results seriously, we need to understand our uncertainties ---
we do not know enough about the Standard Model to make definitive predictions,
but must distribute our unit allotment of probability over various plausible
background predictions.

With such a complicated experiment as \atlas, in a universe with such varied
particles as ours, there are many different factors of uncertainty to consider.
As described in Section~\ref{sec:searches_practice}, the effect of each
factor is studied to assign regularized variations to the Poisson expectations
of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
For the $\twoljets$ search, we used HistFitter to build our every HistFactory
models~\cite{Besjes_2015,baak2015histfitter}.

Thankfully, many standard variations are evaluated automatically with \atlas\
software by simulating alternative samples with varied parameters, then
comparing the resulting histograms.
Such standard variations relate to performance and calibration effects, which
have been carefully studied by many researchers in \atlas, to whom I am deeply
grateful.
They cover variations on our objects --- jets, leptons and $\ptmiss$, as well
as flavour tagging and pileup effects.

Statistical uncertainties on simulated samples are also semi-automatic.
In each bin, its statistical uncertainty is controlled by a single parameter,
which intends to cover the uncertainties on all backgrounds samples
contributing in that bin.
An alternative prescription exists which gives one parameter per sample per
bin, but this greatly increases the complexity of the model and was not
considered necessary.

The $\twoljets$-electroweak search implements a total of $136$ modifiers its
background-only model, and more for each signal model.
Of those I evaluated more manually, some of the more interesting or important
will be discussed in the remainder of this section ---
Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
alternative simulated samples in Section~\ref{sec:2ljets_sample_comparisons},
signal modelling uncertainties in Section~\ref{sec:2ljets_signal_variations}
and theoretical uncertainties in Section~\ref{sec:2ljets_theory_variations}.

@@

\begin{table}[tp]
\centering
\begin{tabular}{lcccccc}
& Prompt (stat)
& MC
& MM
& Mean
& \quad\underline{Max}\hphantom{foo}
\\[0.5em]
CR-Z & $143.6 \pm 4.6$
& $0.1^{+0.2}_{-0.1}$
& $0.01^{+3.2}_{-0.0}$
& $0.02^{+0.14}_{-0.02}$
& $0.1^{+3.2}_{-0.1}$
\\[0.2em]
VR-Int & $18.7 \pm 0.9$
& $0.2^{+0.4}_{-0.2}$
& $0.01^{+0.46}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.2^{+0.5}_{-0.2}$
\\[0.2em]
VR-High & $19.7 \pm 0.7$
& $0.04^{+0.08}_{-0.04}$
& $0.01^{+0.73}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.04^{+0.73}_{-0.04}$
\\[0.2em]
VR-High-1J & $51.9 \pm 1.9$
& $0.003^{+0.005}_{-0.003}$
& $0.01^{+0.59}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.59}_{-0.02}$
\\[0.2em]
SR-Low-b & $15.4 \pm 1.0$
& $0.001^{+0.001}_{-0.001}$
& $0.01^{+0.11}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.17}_{-0.02}$
\\[0.2em]
SR-Int-a & $22.3 \pm 0.8$
& $0^{+0}_{-0}$
& $0.01^{+0.72}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.72}_{-0.02}$
\\[0.2em]
SR-Int-b & $11.4 \pm 0.5$
& $0.002^{+0.003}_{-0.002}$
& $0.01^{+0.43}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.43}_{-0.02}$
\\[0.2em]
SR-High-8-a & $2.4 \pm 0.2$
& $0^{+0}_{-0}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\[0.2em]
SR-High-8-b & $2.4 \pm 0.3$
& $0^{+0}_{-0}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\[0.2em]
SR-High-1J & $1.0 \pm 0.1$
& $0.004^{+0.007}_{-0.004}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\[0.2em]
SR-$\llbb$ & $0.7 \pm 0.1$
& $0^{+0}_{-0}$
& $0.01^{+0}_{-0.01}$
& $0.02^{+0.14}_{-0.02}$
& $0.02^{+0.14}_{-0.02}$
\\
\end{tabular}
\caption[%
Replacement fake estimates with up and down variations
]{%
Replacement fake estimates with up and down variations.
Those variations cover statistical and systematic sources.
The method `mMax' is used for the $\twoljets$-electroweak analysis;
as described in the text, it uses the largest values of central, statistical
and positive systematic variations from other methods, and a down variation
of zero.
}
\label{tab:2ljets_fnp_estimates}
\end{table}

\subsection{Sample comparisons}
\label{sec:2ljets_sample_comparisons}
Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
Figure~\ref{fig:2ljets_low_crz_pre_post}
Figure~\ref{fig:2ljets_vrlow_alt}
Figure~\ref{fig:2ljets_low_vrlow_pre_post}

@@

\subsection{Signals}
\label{sec:2ljets_signal_variations}

% TODO GMSB reweighting
% TODO off-shell plus/minus split
% TODO xsec variations

\subsection{Theory}
\label{sec:2ljets_theory_variations}

% extrapolation
The total theory variation is a combination of seven components representing
\TODO{what are they?}.
A relative uncertainty due to each, $r$, is evaluated at each benchmark
point $\vec y_i$, and extrapolated elsewhere as an exponentially weighted
combination
\begin{equation}
\label{eqn:2ljets_signal_extrap}
r(\vec x) = \frac{
\sum_i r_i e^{-d(\vec x, \vec y_i)}
}{
\sum_i e^{-d(\vec x, \vec y_i)}
}
\end{equation}
in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
physical understanding into the extrapolation.

### 2022-06-27 main/chapter_searches.tex
squeeze significance to fit width
splay out big significance measure a little
add first draft high region def text
begin modelling section; add hist* and pyhf references

\section{Searches}
% histograms, CR VR SR, systematics

@@

\section{Frequentist theory}
\begin{epigraphs}
\qitem{%
Within this theory, statistical methods of great practical usefulness have been
developed, and its statements can and frequently do contribute in a vague way
to the interpretation of data. \ldots%
}%
{John~W.~Pratt,
\textit{Review: Testing Statistical Hypotheses},
1961~\cite{pratt1961testing}}
\end{epigraphs}

\section{Post-frequentist practice}
\begin{epigraphs}
\qitem{%
\ldots\ But this book, by its very excellence, its thoroughness, lucidity and
precision, intensifies my growing feeling that nevertheless the theory is
arbitrary, be it however ``objective,'' and the problems it solves, however
precisely it may solve them, are not even simplified theoretical counterparts
of the real problems to which it is applied.%
}%
{John~W.~Pratt,
\textit{Review: Testing Statistical Hypotheses},
1961~\cite{pratt1961testing}}
\end{epigraphs}

% significancen measures
The ATLAS-recommended~\cite{atlas_significance} significance measure is
\begin{equation}
\label{eqn:significance_atlas}
S_\mathrm{ATLAS}(n, \mu, \sigma)
= \mathrm{sgn}(n - \mu) \sqrt{2\left(
    n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
    - \frac{\mu^2}{\sigma^2}\log\left(
        1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
    \right)
\right)}
~.
\end{equation}

>>>>

\section{Searches}
\label{sec:searches_searches}
% histograms, CR VR SR, systematics

@@

\section{Frequentist theory}
\label{sec:searches_frequentism}
\begin{epigraphs}
\qitem{%
Within this theory, statistical methods of great practical usefulness have been
developed, and its statements can and frequently do contribute in a vague way
to the interpretation of data. \ldots%
}%
{John~W.~Pratt,
\textit{Review: Testing Statistical Hypotheses},
1961~\cite{pratt1961testing}}
\end{epigraphs}

\section{Post-frequentist practice}
\label{sec:searches_practice}
\begin{epigraphs}
\qitem{%
\ldots\ But this book, by its very excellence, its thoroughness, lucidity and
precision, intensifies my growing feeling that nevertheless the theory is
arbitrary, be it however ``objective,'' and the problems it solves, however
precisely it may solve them, are not even simplified theoretical counterparts
of the real problems to which it is applied.%
}%
{John~W.~Pratt,
\textit{Review: Testing Statistical Hypotheses},
1961~\cite{pratt1961testing}}
\end{epigraphs}

% histfactory

To model the consequences of these many factors of uncertainty, each factor is
assigned a real-valued parameter which when varied changed the expected yields
in the hisogram bins of the search.
In this way, we form a statistical model which is a function
\begin{equation}
H_{\!f}(\vec \theta) \rightarrow (\vec \mu, \log f)
\end{equation}
from the parameters $\theta$ to Poisson expectations $\vec\mu$ the histogram
bins and a regularization term $\log f$ which is maximized in fitting.

HistFactory~\cite{cranmer2012histfactory}
pyhf~\cite{heinrich2021pyhf}
HistFitter~\cite{Besjes_2015,baak2015histfitter}

% significancen measures
The ATLAS-recommended significance measure~\cite{atlas_significance} is
\begin{align}
\label{eqn:significance_atlas}
S_\mathrm{\atlas}(n; \mu, \sigma) =~&
\mathrm{sgn}(n - \mu) \times \sqrt{2} \\[0.2em] \nonumber
&
\times \sqrt{
n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
- \frac{\mu^2}{\sigma^2}\log\left(
1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
\right)
}
.
\end{align}
